---
title: "An introduction to BASiCStan"
subtitle: "A Stan implementation of BASiCS"
output: rmarkdown::html_vignette
vignette: >
    %\VignetteIndexEntry{BASiCStan}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

# Outline

This package implements alternative inference methods for
[BASiCS](https://bioconductor.org/packages/release/bioc/html/BASiCS.html).
The original package uses adaptive Metropolis within Gibbs sampling, while
BASiCStan uses Stan to enable the use of maximum *a posterior*, variational
inference, and Hamiltonian Monte Carlo to infer parameters. These each
have advantages for different use cases.


# Use

To use `BASiCStan`, we can first use `BASICS_MockSCE()` to generate an
toy example dataset. We will also set a seed for reproducibility.

```{r setup}
library("BASiCStan")
set.seed(42)
sce <- BASiCS_MockSCE()
```

The interface for running MCMC using BASiCS and using alternative inference
methods using Stan is very similar. It is worth noting that the joint prior
between mean and over-dispersion parameters, corresponding to
`Regression = TRUE`, is the only supported mode in `BASiCStan()`.

```{r basics, results = "hide"}
amwg_fit <- BASiCS_MCMC(
    sce,
    N = 2000,
    Thin = 10,
    Burn = 1000,
    Regression = TRUE
)
stan_fit <- BASiCStan(sce)
```

The output of `BASiCStan()` is an object of class `BASiCS_Chain`, similar to
the output of `BASiCS_MCMC()`. Therefore, . For example, we can plot the
relationship between mean and over-dispersion estimated using the joint
regression prior:

```{r}
BASiCS_ShowFit(amwg_fit)
BASiCS_ShowFit(stan_fit)
```
